---
title: "great_wine_project"
author: "Henry Oamen"
date: "01/06/2024"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
The Great Wine Company

Wine is all about people. We are a company run by wine people. We offer free UK delivery on orders over £500 and 10% off orders of 120+ bottles. We list over 130k wines from over 130 fabulous winemakers across the world in more than 44 countries. We know each of our producers personally. We want to continue to enjoy monopoly and expand into other countries where we are not yet dominant. We now require an analyst to help provide insight into our business. We would like to know which countries produce the finest quality wine (we rate high quality wine as wine with points equal or above 90). Which of our products is the most expensive and which producer makes this. Which products should we not supply anymore, with points less than 45. Please provide a report for the next meeting with the programme director, include some beautiful graphs 


Loading necessary packages

```{r}
library(tidyverse)
```

```{r}
library(readxl)
```
Reading in our data

```{r}
data <- read_csv("winemag-data-130k-v2.csv", na = "")
```

```{r}
view(data)
```
Exploring our data

```{r}
glimpse(data)
```


```{r}
colnames(data)
```

```{r}
head(data)
```


```{r}
#we need to find the total number of rows with na values 
is.na(data) #this returns TRUE values for na-containing cells. 
sum(is.na(data)) # there are many nan in the data
```


```{r}
sum(is.na(data$province))
```

```{r}
data_nan <- data[is.na(data$country), ]

view(data_nan) 
```
Filtering required dataset to exclude NaN values

```{r}
filtered_data <- filter(data, country!='NA' & points!='NA' & price!='NA' & province!='NA' & designation != 'NA') # we can exclude the rows with NaN

view(filtered_data)
```

1. To determine which country produced the finest quality wine based on points

```{r}

country_points <- filtered_data[c('country', 'points', 'price')] # we extract the columns of interest from the filtered data

country_points
```


```{r}
ggplot(country_points, aes(country, points)) + geom_boxplot() + theme_bw() + theme(plot.title = element_text(hjust = 0.5, colour = "steelblue", face = "bold"), axis.text.x = element_text(size = 8, colour = "black", angle = 40, vjust = 1, hjust = 1)) + labs(title = "Wine Points vs Country of production", x = "Country of production", y = "Points")
```

Australia, France, Italy, Portugal and US produce higher quality wine relative to wines from other countries



```{r}

group_data <- country_points %>% group_by(country) %>% slice(which.max(points)) # we select the rows having max points for each country

group_data
```


```{r}
group_data[order(-group_data$points), ] # data sorted in descending order to quickly see the countries with the highest points 
```

We can see that Australia, France, Italy Portugal and US make the best quality wine.

```{r}

quality_wine  <- filter(group_data, points >= 90) # criteria for quality wine is points >= 90, we select countries that meet this criteria for quality wine

quality_wine
```


```{r}
#we can plot the graph of the countries with the highest quality of wine
ggplot(quality_wine, aes(country, points, size = price)) + geom_point() + theme_bw() + theme(plot.title = element_text(hjust = 0.5, colour = "red", face = "bold"), axis.text.x = element_text(size = 10, colour = "black", angle = 40, vjust = 1, hjust = 1)) + labs(title = "Highest points vs Country of production", x = "Country of production", y = "Points")
```


```{r}
cor(country_points$points, country_points$price) # The quality of wine shows some moderate correlation with the price but not a perfect relationship.
```


2. To determine which products is the most expensive and which producer makes this

```{r}
wine_prod <- filtered_data[c('country', 'designation', 'price', 'points')] # we select the desired columns

wine_prod
```

```{r}

wine_prod[order(-wine_prod$price), ] #we sort the data in descending order to find the province producing the most expensive wine
```
The most expensive product is Roger Rose Vineyard which is made in the US.

```{r}
max(wine_prod$price)
```


```{r}
# we can group by country and select only the maximum for each country


wine_prod %>% 
  group_by(country) %>% 
  slice(which.max(price)) %>%
  ggplot(aes(country, price, size = points)) + geom_point() + theme_bw() + theme(plot.title = element_text(hjust = 0.5, colour = "red", face = "bold"), axis.text.x = element_text(size = 9, colour = "black", angle = 40, vjust = 1, hjust = 1)) + labs(title = "Wine Prices vs Country of production", x = "Country of production", y = "Price")

```

US makes the most expensive wine

3. To determine products which should not be supplied anymore based on a benchmark of >=45 points


```{r}
products_filtered <- 
  filtered_data [c('designation', 'points')] %>% 
  filter(points <= 45)
                
products_filtered 


```
There are no products with points less than 45 therefore all products can continue to be supplied.

